---
title: "Machine Learning for HIV expression level prediction"
output: 
  html_document: default
---
by Jeon Lee on October 1, 2019


$~$

## CASE I: numerical ChrommHMM + numerical Lamin
#### Import and preprocess the data 
```{r message=FALSE, warning=FALSE}
library(knitr)
library(kableExtra)
library(dplyr)
library(stringr)
library(dplyr)
library(ggplot2)


setwd("~/R/IvanDorso_lab")
inputData <- read.table("HIVexp_His7Plus4_chrom_lamin_damID_4ML.tsv", sep="\t", header = TRUE) #damID data added

newData <- inputData[, c(6:285)]
newID <- paste(inputData$chr, inputData$start, sep = '_')

# remove duplicate rows
idx_dup <- which(duplicated(newID))
newData <- newData[-idx_dup,]
rownames(newData) <- newID[-idx_dup]

# log10 transform for damID data columns
newData[,c(261:280)] <- data.frame(lapply(newData[,c(261:280)], function(x) {log(x+1e-3,10)}))

# replace '-' sign with 'neg' for column names 
colnames(newData) <- gsub("\\.", "neg", colnames(newData))

# convert ChromHMM and Lamin values into numeric values using lookup tables
tb_ChromHMM <- data.frame(Status=c("U1","U4","U3","U6","U2","U7","U5","U10","U8","U11","U12","U14","U13","U9", "U15"), 
                          Score=c(14:0)/14)
tb_Lamin <- data.frame(Protein=c("A1","A2","B1","B2","B3","B4"), 
                       Score=(c(2,2,-2,-1,-1,0)+2)/4)

newData[,c(1:20)]  <- data.frame(lapply(newData[,c(1:20)],
                                        function(x) {tb_ChromHMM[match(x, tb_ChromHMM$Status), "Score"]}))

newData[,c(21:40)] <- data.frame(lapply(newData[,c(21:40)], function(x) {tb_Lamin[match(x, tb_Lamin$Protein), "Score"]}))

# check data dimension
n_obs <- dim(newData)[1]
n_fea <- dim(newData)[2]

kable(head(newData,10)) %>% kable_styling("striped", full_width = F) %>% scroll_box(width = "100%")
```

$~$

#### HIV expression level histogram 
```{r message=FALSE, warning=FALSE}
newData$HIVexp <- scale(inputData$HIVexp[-idx_dup])
ggplot(newData, aes(x=HIVexp)) + geom_histogram(color="darkblue", fill="lightblue") + labs(title="Histogram of scaled HIV expression levels")
```

$~$

#### HIV expression stratification into Low, Mid and High 
```{r message=FALSE, warning=FALSE}
newData <- newData %>% mutate(ExpLevel = case_when(HIVexp >= 0.5 ~ 1, HIVexp >= -0.5 ~ 0.5, TRUE ~ 0))
kable(data.frame(table(newData$ExpLevel))) %>% kable_styling("striped", full_width = F)

#==== Create Training Data
newData_High <- newData[which(newData$ExpLevel == 1), ]
newData_Mid <- newData[which(newData$ExpLevel == 0.5), ]
newData_Low <- newData[which(newData$ExpLevel == 0), ]
```

$~$

#### Train and test data preparation only for High and Low expressions
```{r message=FALSE, warning=FALSE}
# set up data split ratio = 0.75 (75% of data for traning; 25% of data for test)
set.seed(1234)
splitRatio <- 0.75

#==== Create Training Data
High_train_index <- sample(1:nrow(newData_High), splitRatio*nrow(newData_High))
Mid_train_index <- sample(1:nrow(newData_Mid), splitRatio*nrow(newData_Mid))
Low_train_index <- sample(1:nrow(newData_Low), splitRatio*nrow(newData_Low))

train_High <- newData_High[High_train_index, ]  
train_Mid <- newData_Mid[Mid_train_index, ]
train_Low <- newData_Low[Low_train_index, ]

train_Data <- rbind(train_High, train_Low) #train_Data <- rbind(train_High, train_Mid, train_Low)

#==== Create Test Data
test_High <- newData_High[-High_train_index, ]
test_Mid <- newData_Mid[-Mid_train_index, ]
test_Low <- newData_Low[-Low_train_index, ]

test_Data <- rbind(test_High, test_Low) #test_Data <- rbind(test_High, test_Mid, test_Low)
```

$~$

#### Information Value(IV) calculation 
```{r message=FALSE, warning=FALSE}
library(smbinning)
# segregate continuous and factor variables
factor_vars <- vector()
continuous_vars <- setdiff(colnames(newData), c("HIVexp", "ExpLevel"))
iv_df <- data.frame(VARS=c(factor_vars, continuous_vars), IV=numeric(n_fea))  # init for IV results

# compute IV for categorical vars
for(factor_var in factor_vars){
  smb <- smbinning.factor(train_Data, y="ExpLevel", x=factor_var)  # WOE table
  if(class(smb) != "character"){ # check if some error occured
    iv_df[iv_df$VARS == factor_var, "IV"] <- smb$iv
  }
}
  
# compute IV for continuous vars
for(continuous_var in continuous_vars){
  smb <- smbinning(train_Data, y="ExpLevel", x=continuous_var)  # WOE table
  if(class(smb) != "character"){  # any error while calculating scores.
    iv_df[iv_df$VARS == continuous_var, "IV"] <- smb$iv
  }
}
  
iv_df <- iv_df[order(-iv_df$IV), ]  # sort
iv_df <- within(iv_df, VARS <- factor(VARS, levels=iv_df$VARS))

ggplot(data=iv_df, aes(x=VARS,y=IV)) + 
  geom_bar(position="dodge",stat="identity") + 
  coord_flip() + theme(axis.text.y=element_blank()) +
  ggtitle("Information Values") 

kable(iv_df) %>% kable_styling("striped", full=FALSE) %>%  scroll_box(height = "400px")
```

$~$

#### Filter out variables with low IVs 
```{r message=FALSE, warning=FALSE}
var_notImp <- iv_df$VARS[iv_df$IV<=0.2] # variables not important by Optimal Binning

train_Data2 <- train_Data[,!names(train_Data) %in% var_notImp]
test_Data2 <- test_Data[,!names(train_Data) %in% var_notImp]
```

$~$

#### Build a Logit Model
```{r message=FALSE, warning=FALSE}
pModel <- glm(ExpLevel ~ . -HIVexp, data=train_Data2, family=binomial(link="logit"))
#summary(pModel)
kable(data.frame(pModel$coefficients)) %>% kable_styling("striped", full_width = F) %>%  scroll_box(height = "400px")
```

$~$

#### HIV expression level prediction for the test data 
```{r message=FALSE, warning=FALSE}
predicted <- plogis(predict(pModel, test_Data2))  #or# predicted <- predict(pModel, testData, type="response")

# Decide on optimal prediction probability cutoff for the model
library(InformationValue)
optCutOff <- optimalCutoff(test_Data2$ExpLevel, predicted)[1] 
optCutOff

# Misclassification Error
misClassError(test_Data2$ExpLevel, predicted, threshold = optCutOff)

# ROC
plotROC(test_Data2$ExpLevel, predicted)

# Concordance
kable(data.frame(Concordance(test_Data2$ExpLevel, predicted))) %>% kable_styling("striped", full_width = F)

# Sensitivity and specificity
Sensitivity <- InformationValue::sensitivity(test_Data2$ExpLevel, predicted, threshold = optCutOff)
Specificity <- InformationValue::specificity(test_Data2$ExpLevel, predicted, threshold = optCutOff)
kable(data.frame(cbind(Sensitivity, Specificity))) %>% kable_styling("striped", full_width = F)

# Confusion Matrix
df_CM <- data.frame(InformationValue::confusionMatrix(test_Data2$ExpLevel, predicted, threshold = optCutOff))
colnames(df_CM) <- c("Low", "High")
kable(df_CM) %>% kable_styling("striped", full_width = F)
```

$~$

#### Compare predicted HIV expression values between Low and High groups
```{r message=FALSE, warning=FALSE}
library(ggpubr)
df <- data.frame(cbind(test_Data2$ExpLevel, test_Data2$HIVexp, predicted))
colnames(df) <- c("ExpLevel", "ExpValue", "Predicted")
df$ExpLevel <- ifelse(df$ExpLevel==0, "Low", "High")
df$ExpLevel <- as.factor(df$ExpLevel)

ggplot(df, aes(x=ExpLevel, y=Predicted, fill=ExpLevel)) + 
  geom_boxplot() + geom_jitter(shape=16, position=position_jitter(0.2)) +
  stat_compare_means(label.x = 1.3, label.y = 1.1)
```


$~$

#### (Optional) Check predicted HIV expression values for Mid-group test data
```{r message=FALSE, warning=FALSE}
predicted_Mid <- plogis(predict(pModel, test_Mid))  #or# predicted <- predict(pModel, testData, type="response")
df2 <- data.frame(rbind(cbind(test_Data2$ExpLevel, test_Data2$HIVexp, predicted), cbind(test_Mid$ExpLevel, test_Mid$HIVexp, predicted_Mid)))
colnames(df2) <- c("ExpLevel", "ExpValue", "Predicted")
df2 <- df2 %>% mutate(ExpLevel = case_when(ExpLevel == 1 ~ "High", ExpLevel == 0 ~ "Low", TRUE ~ "Mid"))
df2$ExpLevel <- factor(df2$ExpLevel, ordered = TRUE, levels = c("High", "Mid", "Low"))
ggplot(df2, aes(x=ExpLevel, y=Predicted, fill=ExpLevel)) +
  geom_boxplot() + geom_jitter(shape=16, position=position_jitter(0.2)) +
  stat_compare_means(label.x = 1.3, label.y = 1.1) + scale_fill_manual(values=c("#F8766D", "#00BA38", "#619CFF"))
```

$~$


## CASE II: numerical ChrommHMM + categorical Lamin
#### Import and preprocess the data 
```{r message=FALSE, warning=FALSE}
library(knitr)
library(kableExtra)
library(dplyr)
library(stringr)
library(dplyr)
library(ggplot2)


setwd("~/R/IvanDorso_lab")
inputData <- read.table("HIVexp_His7Plus4_chrom_lamin_damID_4ML.tsv", sep="\t", header = TRUE) #damID data added

newData <- inputData[, c(6:285)]
newID <- paste(inputData$chr, inputData$start, sep = '_')

# remove duplicate rows
idx_dup <- which(duplicated(newID))
newData <- newData[-idx_dup,]
rownames(newData) <- newID[-idx_dup]

# log10 transform for damID data columns
newData[,c(261:280)] <- data.frame(lapply(newData[,c(261:280)], function(x) {log(x+1e-3,10)}))

# replace '-' sign with 'neg' for column names 
colnames(newData) <- gsub("\\.", "neg", colnames(newData))

# convert ChromHMM and Lamin values into numeric values using lookup tables
tb_ChromHMM <- data.frame(Status=c("U1","U4","U3","U6","U2","U7","U5","U10","U8","U11","U12","U14","U13","U9", "U15"), 
                          Score=c(14:0)/14)
tb_Lamin <- data.frame(Protein=c("A1","A2","B1","B2","B3","B4"), 
                       Score=(c(2,2,-2,-1,-1,0)+2)/4)

newData[,c(1:20)]  <- data.frame(lapply(newData[,c(1:20)],
                                        function(x) {tb_ChromHMM[match(x, tb_ChromHMM$Status), "Score"]}))

# check data dimension
n_obs <- dim(newData)[1]
n_fea <- dim(newData)[2]

kable(head(newData,10)) %>% kable_styling("striped", full_width = F) %>% scroll_box(width = "100%")
```

$~$

#### HIV expression level histogram 
```{r message=FALSE, warning=FALSE}
newData$HIVexp <- scale(inputData$HIVexp[-idx_dup])
#ggplot(newData, aes(x=HIVexp)) + geom_histogram(color="darkblue", fill="lightblue") + labs(title="Histogram of scaled HIV expression levels")
```

$~$

#### HIV expression stratification into Low, Mid and High 
```{r message=FALSE, warning=FALSE}
newData <- newData %>% mutate(ExpLevel = case_when(HIVexp >= 0.5 ~ 1, HIVexp >= -0.5 ~ 0.5, TRUE ~ 0))
#table(newData$ExpLevel)

#==== Create Training Data
newData_High <- newData[which(newData$ExpLevel == 1), ]
newData_Mid <- newData[which(newData$ExpLevel == 0.5), ]
newData_Low <- newData[which(newData$ExpLevel == 0), ]
```

$~$

#### Train and test data preparation only for High and Low expressions
```{r message=FALSE, warning=FALSE}
# set up data split ratio = 0.75 (75% of data for traning; 25% of data for test)
set.seed(1234)
splitRatio <- 0.75

#==== Create Training Data
High_train_index <- sample(1:nrow(newData_High), splitRatio*nrow(newData_High))
Mid_train_index <- sample(1:nrow(newData_Mid), splitRatio*nrow(newData_Mid))
Low_train_index <- sample(1:nrow(newData_Low), splitRatio*nrow(newData_Low))

train_High <- newData_High[High_train_index, ]  
train_Mid <- newData_Mid[Mid_train_index, ]
train_Low <- newData_Low[Low_train_index, ]

train_Data <- rbind(train_High, train_Low) #train_Data <- rbind(train_High, train_Mid, train_Low)

#==== Create Test Data
test_High <- newData_High[-High_train_index, ]
test_Mid <- newData_Mid[-Mid_train_index, ]
test_Low <- newData_Low[-Low_train_index, ]

test_Data <- rbind(test_High, test_Low) #test_Data <- rbind(test_High, test_Mid, test_Low)
```

$~$

#### Information Value(IV) calculation 
```{r message=FALSE, warning=FALSE}
library(smbinning)
# segregate continuous and factor variables
factor_vars <- colnames(newData)[21:40]
continuous_vars <- setdiff(colnames(newData)[-c(21:40)], c("HIVexp", "ExpLevel"))
iv_df <- data.frame(VARS=c(factor_vars, continuous_vars), IV=numeric(n_fea))  # init for IV results

# compute IV for categorical vars
for(factor_var in factor_vars){
  smb <- smbinning.factor(train_Data, y="ExpLevel", x=factor_var)  # WOE table
  if(class(smb) != "character"){ # check if some error occured
    iv_df[iv_df$VARS == factor_var, "IV"] <- smb$iv
  }
}
  
# compute IV for continuous vars
for(continuous_var in continuous_vars){
  smb <- smbinning(train_Data, y="ExpLevel", x=continuous_var)  # WOE table
  if(class(smb) != "character"){  # any error while calculating scores.
    iv_df[iv_df$VARS == continuous_var, "IV"] <- smb$iv
  }
}
  
iv_df <- iv_df[order(-iv_df$IV), ]  # sort
iv_df <- within(iv_df, VARS <- factor(VARS, levels=iv_df$VARS))

ggplot(data=iv_df, aes(x=VARS,y=IV)) + 
  geom_bar(position="dodge",stat="identity") + 
  coord_flip() + theme(axis.text.y=element_blank()) +
  ggtitle("Information Values") 

kable(iv_df) %>% kable_styling("striped", full=FALSE) %>%  scroll_box(height = "400px")
```

$~$

#### Filter out variables with low IVs 
```{r message=FALSE, warning=FALSE}
var_notImp <- iv_df$VARS[iv_df$IV<=0.2] # variables not important by Optimal Binning

train_Data2 <- train_Data[,!names(train_Data) %in% var_notImp]
test_Data2 <- test_Data[,!names(train_Data) %in% var_notImp]
```

$~$

#### Build a Logit Model
```{r message=FALSE, warning=FALSE}
pModel <- glm(ExpLevel ~ . -HIVexp, data=train_Data2, family=binomial(link="logit"))
#summary(pModel)
kable(data.frame(pModel$coefficients)) %>% kable_styling("striped", full_width = F) %>%  scroll_box(height = "400px")
```

$~$

#### HIV expression level prediction for the test data 
```{r message=FALSE, warning=FALSE}
predicted <- plogis(predict(pModel, test_Data2))  #or# predicted <- predict(pModel, testData, type="response")

# Decide on optimal prediction probability cutoff for the model
library(InformationValue)
optCutOff <- optimalCutoff(test_Data2$ExpLevel, predicted)[1] 
optCutOff

# Misclassification Error
misClassError(test_Data2$ExpLevel, predicted, threshold = optCutOff)

# ROC
plotROC(test_Data2$ExpLevel, predicted)

# Concordance
kable(data.frame(Concordance(test_Data2$ExpLevel, predicted))) %>% kable_styling("striped", full_width = F)

# Sensitivity and specificity
Sensitivity <- InformationValue::sensitivity(test_Data2$ExpLevel, predicted, threshold = optCutOff)
Specificity <- InformationValue::specificity(test_Data2$ExpLevel, predicted, threshold = optCutOff)
kable(data.frame(cbind(Sensitivity, Specificity))) %>% kable_styling("striped", full_width = F)

# Confusion Matrix
df_CM <- data.frame(InformationValue::confusionMatrix(test_Data2$ExpLevel, predicted, threshold = optCutOff))
colnames(df_CM) <- c("Low", "High")
kable(df_CM) %>% kable_styling("striped", full_width = F)
```

$~$

#### Compare predicted HIV expression values between Low and High groups
```{r message=FALSE, warning=FALSE}
library(ggpubr)
df <- data.frame(cbind(test_Data2$ExpLevel, test_Data2$HIVexp, predicted))
colnames(df) <- c("ExpLevel", "ExpValue", "Predicted")
df$ExpLevel <- ifelse(df$ExpLevel==0, "Low", "High")
df$ExpLevel <- as.factor(df$ExpLevel)

ggplot(df, aes(x=ExpLevel, y=Predicted, fill=ExpLevel)) + 
  geom_boxplot() + geom_jitter(shape=16, position=position_jitter(0.2)) +
  stat_compare_means(label.x = 1.3, label.y = 1.1)
```


$~$

#### (Optional) Check predicted HIV expression values for Mid-group test data
```{r message=FALSE, warning=FALSE}
predicted_Mid <- plogis(predict(pModel, test_Mid))  #or# predicted <- predict(pModel, testData, type="response")
df2 <- data.frame(rbind(cbind(test_Data2$ExpLevel, test_Data2$HIVexp, predicted), cbind(test_Mid$ExpLevel, test_Mid$HIVexp, predicted_Mid)))
colnames(df2) <- c("ExpLevel", "ExpValue", "Predicted")
df2 <- df2 %>% mutate(ExpLevel = case_when(ExpLevel == 1 ~ "High", ExpLevel == 0 ~ "Low", TRUE ~ "Mid"))
df2$ExpLevel <- factor(df2$ExpLevel, ordered = TRUE, levels = c("High", "Mid", "Low"))
ggplot(df2, aes(x=ExpLevel, y=Predicted, fill=ExpLevel)) +
  geom_boxplot() + geom_jitter(shape=16, position=position_jitter(0.2)) +
  stat_compare_means(label.x = 1.3, label.y = 1.1) + scale_fill_manual(values=c("#F8766D", "#00BA38", "#619CFF"))
```

$~$


## CASE III: previous numerical ChrommHMM + categorical Lamin
#### Import and preprocess the data 
```{r message=FALSE, warning=FALSE}
library(knitr)
library(kableExtra)
library(dplyr)
library(stringr)
library(dplyr)
library(ggplot2)


setwd("~/R/IvanDorso_lab")
inputData <- read.table("HIVexp_His7Plus4_chrom_lamin_damID_4ML.tsv", sep="\t", header = TRUE) #damID data added

newData <- inputData[, c(6:285)]
newID <- paste(inputData$chr, inputData$start, sep = '_')

# remove duplicate rows
idx_dup <- which(duplicated(newID))
newData <- newData[-idx_dup,]
rownames(newData) <- newID[-idx_dup]

# log10 transform for damID data columns
newData[,c(261:280)] <- data.frame(lapply(newData[,c(261:280)], function(x) {log(x+1e-3,10)}))

# replace '-' sign with 'neg' for column names 
colnames(newData) <- gsub("\\.", "neg", colnames(newData))

# convert ChromHMM and Lamin values into numeric values using lookup tables
tb_ChromHMM <- data.frame(Status=c("U1","U4","U3","U6","U2","U7","U5","U10","U8","U11","U12","U14","U13","U9", "U15"), 
                          Score=c(14:0)/14)
tb_Lamin <- data.frame(Protein=c("A1","A2","B1","B2","B3","B4"), 
                       Score=(c(2,2,-2,-1,-1,0)+2)/4)

newData[,c(1:20)] <- data.frame(lapply(newData[,c(1:20)], function(x) {as.numeric(as.character(gsub("U", "", x)))}))

# check data dimension
n_obs <- dim(newData)[1]
n_fea <- dim(newData)[2]

kable(head(newData,10)) %>% kable_styling("striped", full_width = F) %>% scroll_box(width = "100%")
```

$~$

#### HIV expression level histogram 
```{r message=FALSE, warning=FALSE}
newData$HIVexp <- scale(inputData$HIVexp[-idx_dup])
#ggplot(newData, aes(x=HIVexp)) + geom_histogram(color="darkblue", fill="lightblue") + labs(title="Histogram of scaled HIV expression levels")
```

$~$

#### HIV expression stratification into Low, Mid and High 
```{r message=FALSE, warning=FALSE}
newData <- newData %>% mutate(ExpLevel = case_when(HIVexp >= 0.5 ~ 1, HIVexp >= -0.5 ~ 0.5, TRUE ~ 0))
#table(newData$ExpLevel)

#==== Create Training Data
newData_High <- newData[which(newData$ExpLevel == 1), ]
newData_Mid <- newData[which(newData$ExpLevel == 0.5), ]
newData_Low <- newData[which(newData$ExpLevel == 0), ]
```

$~$

#### Train and test data preparation only for High and Low expressions
```{r message=FALSE, warning=FALSE}
# set up data split ratio = 0.75 (75% of data for traning; 25% of data for test)
set.seed(1234)
splitRatio <- 0.75

#==== Create Training Data
High_train_index <- sample(1:nrow(newData_High), splitRatio*nrow(newData_High))
Mid_train_index <- sample(1:nrow(newData_Mid), splitRatio*nrow(newData_Mid))
Low_train_index <- sample(1:nrow(newData_Low), splitRatio*nrow(newData_Low))

train_High <- newData_High[High_train_index, ]  
train_Mid <- newData_Mid[Mid_train_index, ]
train_Low <- newData_Low[Low_train_index, ]

train_Data <- rbind(train_High, train_Low) #train_Data <- rbind(train_High, train_Mid, train_Low)

#==== Create Test Data
test_High <- newData_High[-High_train_index, ]
test_Mid <- newData_Mid[-Mid_train_index, ]
test_Low <- newData_Low[-Low_train_index, ]

test_Data <- rbind(test_High, test_Low) #test_Data <- rbind(test_High, test_Mid, test_Low)
```

$~$

#### Information Value(IV) calculation 
```{r message=FALSE, warning=FALSE}
library(smbinning)
# segregate continuous and factor variables
factor_vars <- colnames(newData)[21:40]
continuous_vars <- setdiff(colnames(newData)[-c(21:40)], c("HIVexp", "ExpLevel"))
iv_df <- data.frame(VARS=c(factor_vars, continuous_vars), IV=numeric(n_fea))  # init for IV results

# compute IV for categorical vars
for(factor_var in factor_vars){
  smb <- smbinning.factor(train_Data, y="ExpLevel", x=factor_var)  # WOE table
  if(class(smb) != "character"){ # check if some error occured
    iv_df[iv_df$VARS == factor_var, "IV"] <- smb$iv
  }
}
  
# compute IV for continuous vars
for(continuous_var in continuous_vars){
  smb <- smbinning(train_Data, y="ExpLevel", x=continuous_var)  # WOE table
  if(class(smb) != "character"){  # any error while calculating scores.
    iv_df[iv_df$VARS == continuous_var, "IV"] <- smb$iv
  }
}
  
iv_df <- iv_df[order(-iv_df$IV), ]  # sort
iv_df <- within(iv_df, VARS <- factor(VARS, levels=iv_df$VARS))

ggplot(data=iv_df, aes(x=VARS,y=IV)) + 
  geom_bar(position="dodge",stat="identity") + 
  coord_flip() + theme(axis.text.y=element_blank()) +
  ggtitle("Information Values") 

kable(iv_df) %>% kable_styling("striped", full=FALSE) %>%  scroll_box(height = "400px")
```

$~$

#### Filter out variables with low IVs 
```{r message=FALSE, warning=FALSE}
var_notImp <- iv_df$VARS[iv_df$IV<=0.2] # variables not important by Optimal Binning

train_Data2 <- train_Data[,!names(train_Data) %in% var_notImp]
test_Data2 <- test_Data[,!names(train_Data) %in% var_notImp]
```

$~$

#### Build a Logit Model
```{r message=FALSE, warning=FALSE}
pModel <- glm(ExpLevel ~ . -HIVexp, data=train_Data2, family=binomial(link="logit"))
#summary(pModel)
kable(data.frame(pModel$coefficients)) %>% kable_styling("striped", full_width = F) %>%  scroll_box(height = "400px")
```

$~$

#### HIV expression level prediction for the test data 
```{r message=FALSE, warning=FALSE}
predicted <- plogis(predict(pModel, test_Data2))  #or# predicted <- predict(pModel, testData, type="response")

# Decide on optimal prediction probability cutoff for the model
library(InformationValue)
optCutOff <- optimalCutoff(test_Data2$ExpLevel, predicted)[1] 
optCutOff

# Misclassification Error
misClassError(test_Data2$ExpLevel, predicted, threshold = optCutOff)

# ROC
plotROC(test_Data2$ExpLevel, predicted)

# Concordance
kable(data.frame(Concordance(test_Data2$ExpLevel, predicted))) %>% kable_styling("striped", full_width = F)

# Sensitivity and specificity
Sensitivity <- InformationValue::sensitivity(test_Data2$ExpLevel, predicted, threshold = optCutOff)
Specificity <- InformationValue::specificity(test_Data2$ExpLevel, predicted, threshold = optCutOff)
kable(data.frame(cbind(Sensitivity, Specificity))) %>% kable_styling("striped", full_width = F)

# Confusion Matrix
df_CM <- data.frame(InformationValue::confusionMatrix(test_Data2$ExpLevel, predicted, threshold = optCutOff))
colnames(df_CM) <- c("Low", "High")
kable(df_CM) %>% kable_styling("striped", full_width = F)
```

$~$

#### Compare predicted HIV expression values between Low and High groups
```{r message=FALSE, warning=FALSE}
library(ggpubr)
df <- data.frame(cbind(test_Data2$ExpLevel, test_Data2$HIVexp, predicted))
colnames(df) <- c("ExpLevel", "ExpValue", "Predicted")
df$ExpLevel <- ifelse(df$ExpLevel==0, "Low", "High")
df$ExpLevel <- as.factor(df$ExpLevel)

ggplot(df, aes(x=ExpLevel, y=Predicted, fill=ExpLevel)) + 
  geom_boxplot() + geom_jitter(shape=16, position=position_jitter(0.2)) +
  stat_compare_means(label.x = 1.3, label.y = 1.1)
```


$~$

#### (Optional) Check predicted HIV expression values for Mid-group test data
```{r message=FALSE, warning=FALSE}
predicted_Mid <- plogis(predict(pModel, test_Mid))  #or# predicted <- predict(pModel, testData, type="response")
df2 <- data.frame(rbind(cbind(test_Data2$ExpLevel, test_Data2$HIVexp, predicted), cbind(test_Mid$ExpLevel, test_Mid$HIVexp, predicted_Mid)))
colnames(df2) <- c("ExpLevel", "ExpValue", "Predicted")
df2 <- df2 %>% mutate(ExpLevel = case_when(ExpLevel == 1 ~ "High", ExpLevel == 0 ~ "Low", TRUE ~ "Mid"))
df2$ExpLevel <- factor(df2$ExpLevel, ordered = TRUE, levels = c("High", "Mid", "Low"))
ggplot(df2, aes(x=ExpLevel, y=Predicted, fill=ExpLevel)) +
  geom_boxplot() + geom_jitter(shape=16, position=position_jitter(0.2)) +
  stat_compare_means(label.x = 1.3, label.y = 1.1) + scale_fill_manual(values=c("#F8766D", "#00BA38", "#619CFF"))
```

$~$

### ==End of Report== ###